---
title: "JGR Biogeosciences - Geochemical Figures"
author: "Emily Lacroix\n"
date: "Last update: 1/13/2022"
output: 
  github_document: 
    toc: TRUE
---

# Set-Up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## libraries
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lme4)
library(lmerTest)
library(MuMIn)
library(lmtest)

```

## Data Files

```{r eval=FALSE}
data_file <- "PuuEke_AllData.xlsx"
```


```{r include=FALSE}

data_file <- "~/Box Sync/Stanford (elacroix@stanford.edu)/Research/AnoxicMicrositesHawaii/Data/PuuEke_AllData.xlsx"

```

## Read-in data
```{r}
data <- 
  data_file %>% 
  read_xlsx(sheet = "Data")
```


# Figure 2
## Make tibble 
```{r warning=FALSE}
fig2_data <- 
  data %>% 
  mutate_at(
    vars(horizon), 
    ~ factor(
      .,
      levels = 
        c(
          "Bh1", 
          "Bhg", 
          "Bh2", 
          "Bs", 
          "Bg"
        )
    )
  ) %>% 
  separate(depth, into = c("top", "bottom", "unit"), remove = FALSE) %>% 
  mutate(
    across(bottom, ~if_else(.=="cm", "86", .)),
    across(c(bottom, top), as.numeric),
    avg_depth_cm = (bottom + top) / 2
  ) %>% 
  select(-c(top, bottom, unit))

  
```


## Figure 2a: %C and PFP Distance
```{r}
fig2_data %>% 
  ggplot(aes(x = distance_mm, y = perc_c, shape = horizon)) + 
  geom_line(linetype = 2, alpha = 0.6, size = 0.5) +
  geom_point(size = 3, fill = "white") + 
  scale_x_continuous(
    breaks = seq(from = 0, to = 10, by = 2),
    limits = c(-0.1, 10),
    expand = expansion(mult = c(0, 0.05))
  ) + 
  scale_y_continuous(
    breaks = seq(from = 7, to = 14, by = 1)
  ) + 
  scale_shape_manual(
    values = c(16, 24, 15, 21, 17)
  ) + 
  facet_wrap(
    facets = vars(collection_yr), 
    scales = "free_y",
    ncol = 1
  ) + 
  guides(shape = guide_legend(override.aes = list(size = 4))) + 
  theme_bw() + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.grid = element_blank(),
    aspect.ratio = 1.2,
    legend.position = "right",
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 16),
    axis.title.x = element_text(margin = margin(t = 20)),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  ) + 
  labs(
    x = "Distance from preferential flow path (mm)",
    y = "Carbon (%)",
    shape = NULL
  ) 

```

## Figure 2b: Delta 14C

```{r}
fig2_data %>% 
  ggplot(aes(x = distance_mm, y = c14, shape = horizon)) + 
  geom_line(linetype = 2, alpha = 0.6, size = 0.5) +
  geom_point(size = 3, fill = "white") + 
  scale_x_continuous(
    breaks = seq(from = 0, to = 10, by = 2),
    limits = c(-0.1, 10),
    expand = expansion(mult = c(0, 0.05))
  ) + 
  scale_shape_manual(
    values = c(16, 24, 15, 21, 17)
  ) + 
  facet_wrap(
    facets = vars(collection_yr), 
    scales = "free_y",
    ncol = 1
  ) + 
  expand_limits(y = -90) + 
  scale_y_continuous(breaks = seq(-100, -900, -100)) + 
  guides(shape = guide_legend(override.aes = list(size = 4))) + 
  theme_bw() + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.grid = element_blank(),
    aspect.ratio = 1.2,
    legend.position = "right",
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 16),
    axis.title.x = element_text(margin = margin(t = 20)),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  ) + 
  labs(
    x = "Distance from preferential flow paths (mm)",
    y = expression(paste(Delta^{14}, "C (\u2030)")),
    shape = NULL
  ) 

```

## Figure 2c: C/N
```{r}
fig2_data %>% 
  ggplot(aes(x = distance_mm, y = c_to_n, shape = horizon)) + 
  geom_line(linetype = 2, alpha = 0.6, size = 0.5) +
  geom_point(size = 3, fill = "white") + 
  scale_x_continuous(
    breaks = seq(from = 0, to = 10, by = 2),
    limits = c(-0.1, 10),
    expand = expansion(mult = c(0, 0.05))
  ) + 
  scale_shape_manual(
    values = c(16, 24, 15, 21, 17)
  ) + 
 facet_wrap(
    facets = vars(collection_yr), 
    scales = "free",
    ncol = 1
  ) + 
  guides(shape = guide_legend(override.aes = list(size = 4))) + 
  theme_bw() + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.grid = element_blank(),
    aspect.ratio = 1.2,
    legend.position = "right",
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 16),
    axis.title.x = element_text(margin = margin(t = 20)),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  ) +  
  labs(
    x = "Distance from preferential flow paths (mm)",
    y = "C/N molar ratio",
    shape = NULL
  ) 


```

# Figure 3: Relative change in 14C vs. PFP distance
```{r message=FALSE}

fig3_data <- 
  fig2_data %>% 
  pivot_wider(
    id_cols = c(horizon, collection_yr),
    names_from = distance_mm, 
    values_from = c14
  ) %>% 
  mutate_at(vars(collection_yr), as.factor) %>% 
  mutate(
    reference_14 = `0.5`
  ) %>% 
  mutate(
    across(c(`0.5`, `2`, `4`, `6`, `10`), ~ . - reference_14)
  ) %>% 
  pivot_longer(
    cols = c(`0.5`, `2`, `4`, `6`, `10`),
    names_to = "distance_mm",
    values_to = "relative_change_c14"
  ) %>% 
  mutate_at(vars(distance_mm), as.numeric)

```


```{r}

fig3_data %>% 
  ggplot(aes(x = distance_mm)) + 
  geom_line(
    data = 
      . %>% 
      group_by(distance_mm, collection_yr) %>% 
      summarise(avg_realtive_change_c14 = mean(relative_change_c14)),
    aes(
      y = avg_realtive_change_c14
    ),
    color = "darkgrey",
    linetype = 2,
    size = 0.5
  ) + 
  geom_point(
    data = 
      . %>% 
      group_by(distance_mm, collection_yr) %>% 
      summarise(avg_realtive_change_c14 = mean(relative_change_c14)),
    aes(
      y = avg_realtive_change_c14
    ),
    shape = 8,
    size = 3
  ) + 
  geom_point(
    aes(shape = horizon, y = relative_change_c14), 
    size = 3, 
    fill = "white"
  ) + 
  scale_shape_manual(
    values = c(16, 24, 15, 21, 17)
  ) + 
  scale_x_continuous(
    breaks = seq(from = 0, to = 10, by = 2),
    limits = c(-0.1, 10),
    expand = expansion(mult = c(0, 0.05))
  ) + 
  facet_wrap(facets = vars(collection_yr), ncol = 1) +
  theme_bw() + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.grid = element_blank(),
    aspect.ratio = 1.2,
    legend.position = "right",
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 16),
    axis.title.x = element_text(margin = margin(t = 20)),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  ) +
  labs(
    shape = NULL,
    x = "Distance from preferential flow path (mm)",
    y = expression(paste("Relative Change in ",Delta^{14}, "C (\u2030)"))
  ) 

```


# Figure 4: SRO vs. Distance from PFP

```{r}
fig2_data %>% 
  ggplot(aes(x = distance_mm, y = sro_minerals_mol_kg, shape = horizon)) + 
  geom_line(linetype = 2, alpha = 0.6, size = 0.5) +
  geom_point(size = 3, fill = "white") + 
  scale_x_continuous(
    breaks = seq(from = 0, to = 10, by = 2),
    limits = c(-0.1, 10),
    expand = expansion(mult = c(0, 0.05))
  ) + 
  scale_shape_manual(
    values = c(16, 24, 15, 21, 17)
  ) + 
  facet_wrap(
    facets = vars(collection_yr), 
    scales = "free",
    ncol = 1
  ) + 
  theme_bw() + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.grid = element_blank(),
    aspect.ratio = 1.2,
    legend.position = "right",
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 16),
    axis.title.x = element_text(margin = margin(t = 20)),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  ) + 
  labs(
    x = "Distance from preferential flow path (mm)",
    y = expression("Oxalate extractable Al+Fe+Si (mol "~kg^-1*")"),
    shape = NULL
  ) 


```


# Figure 6: 14C vs. SRO 

```{r message=FALSE}
fig2_data %>% 
  mutate(across(collection_yr, as.factor)) %>% 
  ggplot(aes(x = sro_minerals_mol_kg, y = c14, color = collection_yr)) + 
  geom_smooth(
    method = "lm", 
    se = FALSE,
    color = "darkgray", 
    linetype = "solid"
  ) +
  geom_point(aes(shape = horizon), size = 3) + 
  scale_y_reverse(
    breaks = seq(from = -100, to = -800, by = -100)
  ) +
  scale_shape_manual(
    values = c(16, 2, 15, 1, 17)
  ) + 
  scale_color_manual(
    values = c("black", "darkred"), 
    labels = c("Pedon A", "Pedon B")
  ) +
  expand_limits(y = c(-100, -400)) +
  guides(
    shape = "none"
  ) + 
  theme_bw() + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.grid = element_blank(),
    aspect.ratio = 1.2,
    legend.position = "right",
    axis.text = element_text(size = 13),
    axis.title = element_text(size = 16),
    #axis.title.x = element_text(margin = margin(t = 20)),
    legend.text = element_text(size = 13),
    legend.key.size = unit(0.8, "cm")
  ) + 
  labs(
    y = expression(paste(Delta^{14}, "C (\u2030)")),
    x = expression("Oxalate extractable Al+Fe+Si (mol"~kg^-1*")"),
    color = NULL,
    shape = NULL
  ) 

```

# Table 1: Linear Regression (Models 1-3)
## Model 1 

14C vs. SRO Mineral content
```{r}

sro_14c_lm <- lm(c14~sro_minerals_mol_kg, data = fig2_data)

sro_14c_lm %>% summary()

lmtest::bptest(sro_14c_lm)

```

## Model 2 

14C vs. Depth (cm)
```{r}

depth_14c_lm <- lm(c14~avg_depth_cm, data = fig2_data)

depth_14c_lm %>% summary()

lmtest::bptest(depth_14c_lm)

```

## Model 3

14C vs. distance from PFP
```{r}

dist_14c_lm <- lm(c14~distance_mm, data = fig2_data)

dist_14c_lm %>% summary()

lmtest::bptest(dist_14c_lm)

```

# Table 1: Multiple Linear Regression (Models 4-7)
## Model 4

14C vs. distance + depth

```{r}

dist_depth_14c_lm <- lm(c14~distance_mm + avg_depth_cm, data = fig2_data)

dist_depth_14c_lm %>% summary()

lmtest::bptest(dist_depth_14c_lm)

```

## Model 5

14C vs. distance + SRO

```{r}

dist_sro_14c_lm <- lm(c14~distance_mm + sro_minerals_mol_kg, data = fig2_data)

dist_sro_14c_lm %>% summary()

lmtest::bptest(dist_sro_14c_lm)

```

## Model 6

14C vs. SRO + depth

```{r}

depth_sro_14c_lm <- 
  lm(c14~avg_depth_cm + sro_minerals_mol_kg, data = fig2_data)

depth_sro_14c_lm %>% summary()

lmtest::bptest(depth_sro_14c_lm)

```

## Model 7

14C vs. SRO + depth + distance

```{r}

depth_dist_sro_14c_lm <- 
  lm(c14~avg_depth_cm + sro_minerals_mol_kg + distance_mm, data = fig2_data)

depth_dist_sro_14c_lm %>% summary()

lmtest::bptest(depth_dist_sro_14c_lm)

```


# Table S2: Linear Mixed-Effect Regression (Models 8-14)

## Model 8

```{r}
lmer_c14_sro <- 
  lmer(c14 ~ sro_minerals_mol_kg + (1|collection_yr), data = fig2_data) 

anova(lmer_c14_sro, ddf = "Kenward-Roger")
summary(lmer_c14_sro, ddf = "Kenward-Roger")

r.squaredGLMM(lmer_c14_sro)

```

## Model 9

```{r}
lmer_c14_depth <- 
  lmer(c14 ~ avg_depth_cm + (1|collection_yr), data = fig2_data) 

anova(lmer_c14_depth, ddf = "Kenward-Roger")
summary(lmer_c14_depth, ddf = "Kenward-Roger")

r.squaredGLMM(lmer_c14_depth)

```

## Model 10

```{r}
lmer_c14_dist <- 
  lmer(c14 ~ distance_mm + (1|collection_yr), data = fig2_data) 

anova(lmer_c14_dist, ddf = "Kenward-Roger")
summary(lmer_c14_dist, ddf = "Kenward-Roger")

r.squaredGLMM(lmer_c14_dist)

```

## Model 11

```{r}
lmer_c14_dist_depth <- 
  lmer(c14 ~ distance_mm + avg_depth_cm + (1|collection_yr), data = fig2_data) 

anova(lmer_c14_dist_depth, ddf = "Kenward-Roger")
summary(lmer_c14_dist_depth, ddf = "Kenward-Roger")

r.squaredGLMM(lmer_c14_dist_depth)

```

## Model 12

```{r}
lmer_c14_dist_sro <- 
  lmer(c14 ~ sro_minerals_mol_kg + distance_mm + (1|collection_yr), data = fig2_data) 

anova(lmer_c14_dist_sro, ddf = "Kenward-Roger")
summary(lmer_c14_dist_sro, ddf = "Kenward-Roger")

r.squaredGLMM(lmer_c14_dist_sro)

```


## Model 13

```{r}
lmer_c14_sro_depth <- 
  lmer(c14 ~ sro_minerals_mol_kg + avg_depth_cm + (1|collection_yr), data = fig2_data) 

anova(lmer_c14_sro_depth, ddf = "Kenward-Roger")
summary(lmer_c14_sro_depth, ddf = "Kenward-Roger")

r.squaredGLMM(lmer_c14_sro_depth)

```

## Model 14

```{r}
lmer_c14_sro_depth_dist <- 
  lmer(c14 ~ 
         sro_minerals_mol_kg + avg_depth_cm + distance_mm + (1|collection_yr), 
       data = fig2_data
    ) 

anova(lmer_c14_sro_depth_dist, ddf = "Kenward-Roger")
summary(lmer_c14_sro_depth_dist, ddf = "Kenward-Roger")

r.squaredGLMM(lmer_c14_sro_depth_dist)

```

# Figure S1
```{r}
fig2_data %>% 
  ggplot(aes(x = distance_mm, y = perc_n, shape = horizon)) + 
  geom_line(linetype = 2, alpha = 0.6, size = 0.5) +
  geom_point(size = 2) + 
  scale_x_continuous(
    breaks = seq(from = 0, to = 10, by = 2),
    limits = c(-0.1, 10),
    expand = expansion(mult = c(0, 0.05))
  ) + 
  scale_shape_manual(
    values = c(16, 2, 15, 1, 17)
  ) + 
 facet_wrap(
    facets = vars(collection_yr), 
    scales = "free",
    ncol = 1
  ) + 
  theme_bw() + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    panel.grid = element_blank(),
    aspect.ratio = 1.2,
    legend.position = "right",
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14)
  ) + 
  labs(
    x = "Distance from preferential flow paths (mm)",
    y = "Nitrogen (%)",
    shape = NULL
  ) 
```

# Figure S7: Residual plots 

## Residuals of Model 1: 14C vs. SRO Mineral content 

```{r}

fig2_data %>% 
  select(c14, sro_minerals_mol_kg, horizon, collection_yr) %>% 
  mutate(
    predicted = predict(sro_14c_lm), 
    residuals = residuals(sro_14c_lm),
    across(collection_yr, as_factor)
  ) %>% 
  ggplot(aes(x = sro_minerals_mol_kg)) + 
  geom_hline(aes(yintercept = 0), color = "darkgray", linetype = 2) +
  geom_point(
    aes(y = residuals, shape = horizon, color = collection_yr), 
    size = 3
  ) +
  scale_color_manual(
    values = c("black", "darkred"), 
    labels = c("Pedon A", "Pedon B")
  ) +
  scale_y_continuous(limits = c(-100, 100)) +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16)
  ) + 
  labs(
    x = expression("Oxalate extractable Al+Fe+Si (mol"~kg^-1*")"),
    y = "Model residuals",
    shape = "Horizon",
    color = NULL
  )

```

## Residuals of Model 5: 14C vs. SRO + Distance

```{r}

fig2_data %>% 
  select(c14, sro_minerals_mol_kg, horizon, distance_mm, collection_yr) %>% 
  mutate(
    predicted = predict(dist_sro_14c_lm), 
    residuals = residuals(dist_sro_14c_lm),
    across(collection_yr, as_factor)
  ) %>% 
  ggplot(aes(x = sro_minerals_mol_kg)) + 
  geom_hline(aes(yintercept = 0), color = "darkgray", linetype = 2) +
  geom_point(aes(y = residuals, shape = horizon, color = collection_yr), size = 3) +
  scale_color_manual(
    values = c("black", "darkred"), 
    labels = c("Pedon A", "Pedon B")
  ) +
  scale_y_continuous(limits = c(-100, 100)) +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16)
  ) + 
  labs(
    x = expression("Oxalate extractable Al+Fe+Si (mol"~kg^-1*")"),
    y = "Model residuals",
    shape = "Horizon",
    color = NULL
  )

```
Residuals appear less patterned and overall smaller. 

## Residuals of Model 6: 14C vs. SRO + depth

```{r}

fig2_data %>% 
  select(c14, sro_minerals_mol_kg, horizon, distance_mm, collection_yr) %>% 
  mutate(
    predicted = predict(depth_sro_14c_lm), 
    residuals = residuals(depth_sro_14c_lm),
    across(collection_yr, as_factor)
  ) %>% 
  ggplot(aes(x = sro_minerals_mol_kg)) + 
  geom_hline(aes(yintercept = 0), color = "darkgray", linetype = 2) +
  geom_point(aes(y = residuals, shape = horizon, color = collection_yr), size = 3) +
  scale_color_manual(
    values = c("black", "darkred"), 
    labels = c("Pedon A", "Pedon B")
  ) +
  scale_y_continuous(limits = c(-100, 100)) +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16)
  ) + 
  labs(
    x = expression("Oxalate extractable Al+Fe+Si (mol"~kg^-1*")"),
    y = "Model residuals",
    shape = "Horizon",
    color = NULL
  )

```


## Residuals of Model 7: 14C vs. SRO + depth + distance

```{r}

fig2_data %>% 
  select(c14, sro_minerals_mol_kg, horizon, distance_mm, collection_yr) %>% 
  mutate(
    predicted = predict(depth_dist_sro_14c_lm), 
    residuals = residuals(depth_dist_sro_14c_lm),
    across(collection_yr, as_factor)
  ) %>% 
  ggplot(aes(x = sro_minerals_mol_kg)) + 
  geom_hline(aes(yintercept = 0), color = "darkgray", linetype = 2) +
  geom_point(aes(y = residuals, shape = horizon, color = collection_yr), size = 3) +
  scale_color_manual(
    values = c("black", "darkred"), 
    labels = c("Pedon A", "Pedon B")
  ) +
  scale_y_continuous(limits = c(-100, 100)) +
  theme_bw() + 
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 20),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 16)
  ) + 
  labs(
    x = expression("Oxalate extractable Al+Fe+Si (mol"~kg^-1*")"),
    y = "Model residuals",
    shape = "Horizon",
    color = NULL
  )

```




